Spatial Reasoning and Statistical Graphics

### 2014-10-11 Paper/Draft1/TestingVisualAptitude.Rnw
Remove comment about preliminary results as the comment no longer applies.
Rearrange paper text and comment out comments that are no longer new/relevant.

% \comment{The text makes still a lot of references to `preliminary results'. Don't just use copy \& paste for the changes!!!! Go through those sections and adjust. We are working on the final analysis now. Even if numbers might still change, they are not preliminary anymore. }
\section{Introduction}
Relevant literature: 
\begin{itemize}
\item \citet{shah1995conceptual} showed that spatial ability was not correlated with accuracy on a simple two-dimensional line graph description task, but that mathematical ability was correlated with accuracy. 
\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same. 
\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
\end{itemize}

@@

\comment{\textit{We might need to expand this a bit to account for the figure classification section, which has some questions with 2 answers and some with 3. Or we could leave it as an exercise for the reader :)}
It's fairly trivial to apply this to the mixture - the only assumption that goes into the variance discussion is that the questions are independent. We don't need to include it here, but you should apply it to the figure classification set.

}

\comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}

\comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown. It would be good to scale everything, so that we have the same theoretical range. That will actually allow a comparison across different tests.}
<<means>>=
library(plyr)
ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
@

\comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population). }

\comment{
Split the next sentence up into two paragraph: what is shown in each of the figures, what are the immediate conclusions? chuck out the gaming hours from figure \ref{fig:visualaptitudects}, that's a demograph, so move it to the other plot (I am aware that nine demographics are easier to display than ten, but content trumps). \\
\textit{Actually, they were split up like that because of continuous vs. categorical. It's not easy to facet when you have different $x$ scales. I've removed video games entirely for the time being. I agree it didn't really belong there, but I can't get it to work with the facetting, either. I will probably just bin it eventually. 
}}

% \comment{What happened to the arts skills? I seem to have a positive association now - why is my data different from yours? \\
% \textit{I have the same updated results you do. I see that Stephanie changed the data file (but didn't add any more rows)... is it possible something changed there? Otherwise, maybe my dropbox didn't update correctly the first time? Or maybe it was figure caching?}}
Results are presented graphically in figures \ref{fig:visualaptitudects} and \ref{fig:visualaptitudecat}.

@@

\comment{\textit{I did change the lineup scoring - for some reason, I was averaging before (and I hadn't included the third battery of lineups b/c of scoring issues). That explains the scale, but not the change in art skills association.}}

@@

Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test. There is little difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class. There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}. In addition, completion of Calculus I is associated with increased performance on lineups (though completion of calculus is also associated with sex). AutoCAD experience is also not significantly associated with lineup performance; there is a difference in the medians, but it does not rise to the level of significance. There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and may be at least partially a result of the large difference in performance due to sex.

>>>>

\section{Introduction}
Relevant literature: 
\begin{itemize}
\item \citet{shah1995conceptual} showed that spatial ability was not correlated with accuracy on a simple two-dimensional line graph description task, but that mathematical ability was correlated with accuracy. 
\item \citet{just1985cognitive} showed that high-spatial-ability viewers used different rotation strategies than low-spatial-ability viewers when asked to whether three-dimensional alphabet cubes were the same. 
\item \citet{hofmann2012graphical} for lineup stimuli and general lineup performance
\end{itemize}

@@

% \comment{\textit{So are we scaling things by the variance/std dev. under random guessing? Or by the theoretical test range?}\\
% Let's use the standard deviation, so that all tests have an expected value of 0 and a variance/standard deviation of 1 under random guessing.}
% 
% \comment{The next step is an overview of the possible ranges of the test scores in our study, because that is driven by the number of tests shown. It would be good to scale everything, so that we have the same theoretical range. That will actually allow a comparison across different tests.}
<<means>>=
library(plyr)
ldply(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")], c(mean=mean,sd= sd))
@

\comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population). \textit{I'm not sure where to find the ``normal scores" in the factor-referenced manual. I'll check through it again, but I don't recall seeing it.}}

@@

Figure \ref{fig:visualaptitudecat} shows participants' responses to the questionnaire given at the beginning of the study; these demographic questions allow us to compare the participants in our study to the undergraduate population of Iowa State as well as to explore relationships between demographic characteristics (major, research experience, etc.) and score on various sections of this test.

@@

There is no significant difference in lineup performance for participants of different age, self-assessed skill rating, previous participation in math or science research or completion of a statistics class. There is a significant difference between male and female performance on lineups; this is not particularly surprising, since men perform better on many spatial tests \citep{voyer1995magnitude} and performance on spatial tests is correlated with phase of the menstrual cycle in women \citep{hausmann2000sex}. In addition, completion of Calculus I is associated with increased performance on lineups (though completion of calculus is also associated with sex). AutoCAD experience is also not significantly associated with lineup performance; there is a difference in the medians, but it does not rise to the level of significance. There is also a significant association between hours of video games played per week and score on lineups, however, this association is not monotonic and may be at least partially a result of the large difference in performance due to sex.
